Lightweight Scheduling for the PRAGMA Cloud Testbed

### 2016-10-17 paper.tex
Add Beth and Renato's changes

\begin{abstract}
The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas such as biodiversity distribution and lake ecology.  PRAGMA's members collaborate on a testbed infrastructure and over the past four years, the technology focus has shifted to cloud and software defined networking as enabling technologies.  This paper describes the design of a web-based cloud scheduler reservation system that enables users to easily run and manage virtual clusters for their science.  Based on a client-server architecture, the cloud scheduler was designed so that it requires minimal installation and and management effort for participating PRAGMA testbed sites.   The PRAGMA cloud scheduler is built using a web-based room reservation tool called Booked and leverages several PRAGMA technologies such as pragma\_boot, Personal Cloud Controller,  and virtual network overlays (e.g., IPOP, ViNe).  We discuss our implementation of the pilot cloud scheduler and then describe future work.   
%add contribute and leverage shared resources
\end{abstract}

@@

The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) researchers actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.  Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.  This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and to conduct scientific experiments.  Testbed sites needed to install at a minimum the Globus toolkit~\cite{globus}, a local HPC job scheduler as well as other optional software such as MPICH-G2~\cite{mpichg2} and Ninf-G~\cite{ninfg}.  In November 2006, nineteen sites in thirteen countries were part of the testbed.  By 2009, the testbed had grown to twenty-seven sites in fifteen countries with a total of 1008 CPUs, more than 1.3 terabytes of memory, and over 24.7 terabytes of online storage.  However, as of 2011, the number of sites started to decline and maintaining the numerous and complex middleware and scientific applications at each local site required significant people effort and expertise.  Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and to lower the amount of effort any site would need to participate in the testbed.

The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
One early example in 2011 was an AutoDock virtual cluster created for the Avian Flu Grid research team~\cite{pragmaReport2011}.   By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory and 657 terabytes of online storage.  To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus in 2013 to the creation and management of virtual clusters.  The following technologies have been explored to facilitate the operation and use of virtual clusters on the PRAGMA cloud testbed.

@@

\textbf{Software-Defined Networking}:   PRAGMA began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network among multi-site virtual clusters and to protect access to sensitive datasets.  PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.  In 2014, The PRAGMA-ENT team worked to create a reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.

@@

PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users to access the resources at a  given time:

\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it had to be lightweight and required  minimal effort and minimal expertise for a site to add their cloud deployment to the list of resources available for scheduling.   Similarly, the scheduler would need to work across multiple cloud systems (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install additional cloud deployment tools.

\textbf{Easy to use}:  Currently, to execute a virtual cluster on the PRAGMA cloud testbed, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and launch their virtual cluster either using pragma\_boot  if it is available or manually via a sequence of defined steps.   The design of the PRAGMA cloud scheduler should first simplify this process by ensuring that pragma\_boot gets deployed more broadly and port it to more cloud systems as needed.   The scheduler should also provide a simple web interface for users to see the available resources and sites, construct their virtual cluster, and manage their images.

@@

This paper described the the design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.  Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation and maintenance overhead and opted for a simple client-server design.  The server has an intuitive web GUI frontend based on the Booked web reservation system software.

In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.  
Currently, the virtual cluster startup latency from the reservation to its availability to the user has not been analyzed in detail but we observed
it is on the order of tens of minutes.   Multiple factors influence this latency, such as the underlying virtualization manager and load on the
physical system where the virtual cluster is instantiated. There is additional latency if the cluster image is not available on the site as it will
need to be downloaded from a remote repository and processed to make it available on the site. While this applies only to the first virtual cluster instantiation, it can add a considerable time  latency  depending on the cluster image size and the network speed from the site to the remote repository.

In the future when the cloud scheduler is deployed to more sites, we can gather statistics on its usability and user experiences with virtual
clusters startup latencies in addition to its applicability to different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the
cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it is used in personal mode allowing us to keep a light footprint on
each of the PRAGMA sites.  Finally, we will investigate using the remote repository option in pragma\_boot to manage application-specific  virtual cluster
images and staging them to the sites.     We will package and document the resulting software so that new users can quickly start using the PRAGMA cloud testbed as an experimental service to test new science and infrastructure tools.

>>>>

\begin{abstract}
The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas such as biodiversity distribution and lake ecology.  PRAGMA's members collaborate on a testbed infrastructure and over the past four years, the technology focus has shifted to cloud and software defined networking as enabling technologies.  This paper describes the design of a web-based cloud scheduler reservation system that enables users to easily run and manage virtual clusters for their science.  Based on a client-server architecture, the cloud scheduler is designed to require minimal installation and management effort for participating PRAGMA testbed sites.   
%remove duplicate "and"; make present tense, simple substitution of enables
The PRAGMA cloud scheduler is built using a web-based room reservation tool called Booked and leverages several PRAGMA technologies such as pragma\_boot, Personal Cloud Controller,  and virtual network overlays (e.g., IPOP, ViNe).  We discuss our implementation of the pilot cloud scheduler and then describe future work.   
%add contribute and leverage shared resources
\end{abstract}

@@

The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) researchers actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.  Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.  This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and to conduct scientific experiments.  Testbed sites needed to install at a minimum the Globus toolkit~\cite{globus}, a local HPC job scheduler as well as other optional software such as MPICH-G2~\cite{mpichg2} and Ninf-G~\cite{ninfg}.  In November 2006, nineteen sites in thirteen countries were part of the testbed.  By 2009, the testbed had grown to twenty-seven sites in fifteen countries with a total of 1008 CPUs, more than 1.3 terabytes of memory, and over 24.7 terabytes of online storage.  However, as of 2011, the number of sites started to decline and maintaining the numerous and complex middleware and scientific applications at each local site required significant people effort and expertise.  Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and to lower the amount of effort any site would need to expend to 
%add "expend to"
participate in the testbed.

The first phase of the PRAGMA cloud testbed started with three sites and focused on
%replace "around" with "on"
 the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
One early example in 2011 was an AutoDock virtual cluster created for the Avian Flu Grid research team~\cite{pragmaReport2011}.   By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory and 657 terabytes of online storage.  To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus in 2013 to the creation and management of virtual clusters.  The following technologies have been explored to facilitate the operation and use of virtual clusters on the PRAGMA cloud testbed.

@@

\textbf{Software-Defined Networking}:   PRAGMA community members
%add "community members"
 began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network among multi-site virtual clusters and to protect access to sensitive datasets.  PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.  In 2014, the PRAGMA-ENT team worked to create a reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.

@@

The PRAGMA resource providers
%added "resource providers"
 had the following three key requirements for a cloud testbed scheduler to enable multiple users to access the resources at a  given time:

\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it 
%remove "had to"; change to present tense
be lightweight and require  minimal effort and minimal expertise for a site to add their cloud deployment to the list of resources available for scheduling.   Similarly, the scheduler would need to work across multiple cloud systems (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install additional cloud deployment tools.

\textbf{Easy to use}:  Currently, to execute a virtual cluster on the PRAGMA cloud testbed, users must
%change "needs to" to "must"
 contact each site individually to gain access to their cloud deployment, upload their images, and launch their virtual cluster either using pragma\_boot  if it is available or manually via a sequence of defined steps.   The design of the PRAGMA cloud scheduler should first simplify this process by ensuring that pragma\_boot gets deployed more broadly and port it to more cloud systems as needed.   The scheduler should also provide a simple web interface for users to see the available resources and sites, construct their virtual cluster, and manage their images.

@@

This paper describes the
%present tense and remove extra "the"
design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.  Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use, low installation and maintenance overhead, and thus opted for a simple client-server design. 
%clarify what was prioritized
The server has an intuitive web GUI front-end based on the Booked web reservation system software.

In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address these.
%these instead of them  
Currently, the virtual cluster startup latency from the point in time of its reservation to its availability to the user has not been analyzed in detail but we observed
it to be on the order of tens of minutes.   
%clarify latency applied to
Multiple factors influence this latency, such as the underlying virtualization manager and load on the
physical system where the virtual cluster is instantiated. There is additional latency if the cluster image is not available on the site as it will
need to be downloaded from a remote repository and processed to make it available on the site. While this applies only to the first virtual cluster instantiation, it can add a considerable time  latency  depending on the cluster image size and the network speed from the site to the remote repository.

Once
%"once" instead of "in the future"
 the cloud scheduler is deployed to more sites, we can gather statistics on its usability and user experiences with virtual
clusters startup latencies in addition to its applicability to different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the
cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it is used in personal mode allowing us to keep a light footprint at
%"at" instead of "on"
each of the PRAGMA sites.  Finally, we will investigate using the remote repository option in pragma\_boot to manage application-specific  virtual cluster
images and staging them to the sites.     We will package and document the resulting software so that new users can quickly start using the PRAGMA cloud testbed as an experimental service to test new science and infrastructure tools.
